---
title: "P8105_hw3_xw2757"
output: github_document
---
```{r}
library(tidyverse)
library(viridis)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

##Probem 1 
```{r}
library(p8105.datasets)
data("instacart")
```

**Inspect data**
```{r}
summary(instacart) 
```
^no missing values


**description of data**
The "insatcart" data set has a total of `r nrow(instacart)` observations of `r ncol(instacart)` variables, including `r names(instacart)` . There are a total of `r count(distinct(instacart, order_id))` orders, and `count(distinct(instacart, user_id))` users. Detailed Data structures are as described below:
```{r}
str(instacart)
```

**1. How many aisles are there, and which aisles are the most items ordered from?**
```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_items = n()) %>% 
  arrange(desc(n_items))
```
There are a total of `r count(distinct(instacart, aisle))` aisles, the most items are ordered from "fresh vegetables" aisle. 


**Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.**

```{r}
instacart %>% 
  group_by(aisle) %>% 
  filter(n() > 10000) %>% 
  summarize(n_items = n()) %>%
  ggplot(aes(x = aisle, y = n_items)) +
  geom_bar(stat = "identity") + coord_flip() + 
  theme(axis.text.y = element_text(size = 6, hjust = 1, vjust = 0.5))
```



**Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.**

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_buy=n()) %>% 
  filter(rank(desc(n_buy))<=3) %>% 
  knitr::kable()
```
The most popular item in the baking ingredients aisle are light brown sugar, organic vanilla extract, and pure baking soda. 
The most popular item in the dog food care aisle are Organix Chicken & Brown Rice Recipe, Organix Grain Free Chicken & Vegetable Dog Food, and Original Dry Dog.
The most popular item in thepackaged vegetables fruits aisle are Organic Baby Spinach, Organic Blueberries, and Organic Raspberries. Number of items bought are shown in the table. 

**Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).**
```{r}
instacart %>% 
  mutate(t_day_number = order_dow + 1) %>% 
  mutate(order_dow  = lubridate::wday(t_day_number, label = TRUE)) %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hr = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hr)
```
Table above shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. 

##Problem 2
```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

**Data cleaning**
```{r}
brfss_clean <- 
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health",
         response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) %>% 
  mutate(response = forcats::fct_relevel(response, c("Poor", "Fair", "Good", "Very good", "Excellent")))
```
**In 2002, which states were observed at 7 or more locations? What about in 2010?**

```{r}
brfss_clean %>% 
  filter(year == "2002") %>% 
  group_by(locationabbr) %>% 
  summarize(n_locations = n_distinct(locationdesc)) %>% 
  filter(n_locations >= 7) %>% 
  knitr::kable()

brfss_clean %>% 
  filter(year == "2010") %>% 
  group_by(locationabbr) %>% 
  summarize(n_locations = n_distinct(locationdesc)) %>% 
  filter(n_locations >= 7) %>% 
  knitr::kable()
```
In 2002, CT, FL, MA, NC, NJ, PA were observed at 7 or more locations. 
In 2010, CA, CO, Fl, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, and WA were observed at 7 or more locations. 
  

**Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. **


```{r}
brfss_01 <- 
  brfss_clean %>% 
  group_by(year, locationabbr) %>% 
  filter(response == "Excellent") %>% 
  summarize(avg_val = mean(data_value), na.rm = TRUE)
```

**Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).**

```{r}
brfss_01 %>% 
  ggplot(aes(x = year, y = avg_val, group=locationabbr, color = locationabbr)) +
  geom_line()+
  theme(legend.text = element_text(size = 4),
        legend.title = element_text(size = 4),
        legend.position = "right",
        legend.box = "horizontal")
```


**Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.**
```{r}
data_response_NY_06_10 <- 
   brfss_clean %>% 
   filter(year %in% c("2006", "2010"), locationabbr == "NY") %>% 
  group_by(locationdesc, response) %>% 
  ggplot(aes(x = data_value, fill = response)) +
  geom_histogram(position = "dodge", binwidth = 8) +
  facet_grid(~year) +
  labs(
    title = "  Distribution of data_value for responses in NY"
  )
  
```


##Problem 3

**Loading data**
```{r}
accelerometer <- 
  read_csv("accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(wknd = ifelse(day %in% c("Saturdat", "Sunday"), "Weekend", "Weekday")) %>% 
  pivot_longer(activity_1:activity_1440, names_to = "minute", names_prefix = "activity_", values_to = "activity") %>% 
  select(wknd, everything()) %>% 
  mutate(minute = as.numeric(minute))
```

**Describe the resulting dataset (e.g. what variables exist, how many observations, etc).**
The resulting data set "activity" contains the following variables: `r names(accelerometer)`, with a total of `r nrow(accelerometer)` observation. 

**Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?**

```{r}
accelerometer %>% 
  group_by(wknd, day_id) %>% 
  summarize(total_activity_day = sum(activity)) %>% 
  knitr::kable()

accelerometer %>% 
  group_by(wknd, day_id) %>% 
  summarize(total_activity_day = sum(activity)) %>% 
  ggplot(aes(x = day_id, y = total_activity_day)) +
  geom_line()

accelerometer %>% 
  group_by(wknd, day_id) %>% 
  summarize(total_activity_day = sum(activity)) %>% 
  ggplot(aes(x = day_id, y = total_activity_day, color = wknd)) +
  geom_line()
```
^can't find apparent trend of overall total day activity. Weekday total activity fluctuates but weekend total activity appears to resembles a downward trend. 

**Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.**

```{r}
accelerometer %>% 
  group_by(day, day_id) %>% 
  ggplot(aes(x = minute, y = activity, group = day_id, color = day)) +
  geom_line(alpha = 0.5)
```
The plot shows that the subject is less active during nighttime, more activities during the day, and the most active time tends to be evening, and tends to be on Fridays. 


